---
title: "CCT Hierarchy Functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CCT Hierarchy Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Load from Cache, Update if New Lineages Found or Create Cache if Missing

- Cache is written to a basic 'rda' file in the user's local AppData directory
- Will not overwrite cache if update has no new lineages compared to cache
- The force argument forcibly returns the lineages in the file specified regardless of how it compares to the cache
- Can supply either `lineages_url` or `aliases_url` to target specific files with `updateHierarchies()`
- checkCache returns a table of lineage hierarchies by default or a data.tree network with `network = TRUE`

### Table

```{r}
library(CCT)
# Grab an older version
cachedat <- checkCache(lineages_url = "https://raw.githubusercontent.com/cov-lineages/pango-designation/4b5957e151c10e92cb49e9df904c0b2ea841f3eb/lineage_notes.txt")

head(cachedat)
```


```{r}
# how many lineages?
nrow(cachedat)
```


```{r}
# Grab latest version
cachedat <- checkCache()

# how many lineages?
nrow(cachedat)
```

### Network
```{r}
cachedatNET <- checkCache(network = TRUE)
```


## Run All in One Line

```{r}
# In case newest version breaks use: lineages_url = "https://raw.githubusercontent.com/TheZetner/pango-designation/master/lineage_notes.txt"

hierarchy <-
  CCT::updateHierarchies() %>% CCT::fromDFtoNetwork()

hierarchy
```

## Update Hierarchies

- Lineages: https://github.com/cov-lineages/pango-designation/blob/master/lineage_notes.txt
- Aliases: https://github.com/cov-lineages/pango-designation/blob/master/pango_designation/alias_key.json
- Can supply a specific local file or version of the files above
- Reports which file failed if one does
- Returns a table of Parent/Lineage/Children
```{r}
l <- CCT::updateHierarchies(lineages_url = "https://raw.githubusercontent.com/TheZetner/pango-designation/master/lineage_notes.txt")

head(l)
```

## From Dataframe to Network
- Take the dataframe from `updateHierarchies()` or `checkCache(network = FALSE)` and build a node network using [data.tree](https://cran.r-project.org/web/packages/data.tree/vignettes/data.tree.html)
- Defines the parent-child relationship between each node (lineage)
- Also returns a table of "orphan" lineages that are missing parents in the table as these break the network and are removed
- Can replace `ll$network` with `cachedatNET` from above
```{r}
ll <- fromDFtoNetwork(l)
ll
```

## Pull all the Ancestors of a Node

### Network
```{r}
ancestors <- pullAncestors(ll$network, "BA.5.2", table = FALSE)
# ancestors <- pullAncestors(cachedatNET, "BA.5.2", table = FALSE)

ancestors
```

### Table
```{r}
ancestors <- pullAncestors(ll$network, "BA.5.2", table = TRUE)
ancestors
```

### Vector
```{r}
# vector supersedes table param
ancestors <- pullAncestors(ll$network, "BA.5.2", table = TRUE, vector = TRUE)
ancestors
```

## Get all the Descendants of a Node

### Network
```{r}
getDescendants(ll$network, "BF.7", table = FALSE) %>% plot()
```

### Table
```{r}
getDescendants(ll$network, "BF.7", table = TRUE)
```

### Vector
```{r}
# vector supersedes table param
getDescendants(ll$network, "BF.7", table = TRUE, vector = TRUE)
```

## Make YAML Watchlist (for Augur)
```{r}
getDescendants(ll$network, "BF.7", vector = TRUE) %>% makeWatchlist()
```

## Download Variants table from CoVariants
```{r}
vars <- getVariants()
vars
```

## Expand Variants Table to Include all Children
```{r}
expandChildren(head(vars), ll$network)
```

# Custom Hierarchies
- CCT can also examine custom hierarchies for descendants and ancestors
- This capacity is limited at the moment but can easily provide the same functionality as above to work with yaml-defined hierarchies.

## Create a Randomly Nested YAML
```{r ch1}
# Random YAML
yaml <- "fruit:\n"
depth <- ""
maxdepth <- 1
deepest <- stringr::fruit[1]
for(i in stringr::fruit){
  flip <- round(runif(1, 0, 1))
  if(flip == 1){
    flip <- round(runif(1, 0, 1))
    if(flip == 1){
      depth <- paste0(depth, "  ")
      if(stringr::str_count(depth, "  ") > maxdepth){
        maxdepth <- stringr::str_count(depth, "  ")
        deepest <- i
      }
    }
  }else{
    depth <- "  "
  }
  yaml <- paste0(yaml,
                 paste0(depth, i, ":\n"))
}
cat(yaml)
```

Now that the hierarchy is defined in the above yaml text we can load it and convert it into a data.tree node network. A YAML file can also be supplied using the commented text.

```{r ch2}

# To load a file
# yaml::read_yaml(yaml) %>%
#   data.tree::as.Node(interpretNullAsList = TRUE)

hier <- yaml::yaml.load(yaml) %>%
  data.tree::as.Node(interpretNullAsList = TRUE)

str(hier)

```

The deepest element of that fake yaml is **`r deepest`** let's use that as an example.

## Examining the Custom Hierarchy

Pull ancestors of deepest and return a node.
```{r ch3}
y <- pullAncestors(x = hier, 
              node_name = deepest)

str(y)
```

Print it (looks like a table)
```{r ch4}
y
```

Return an actual table. Note the lineage column of `NA`.
```{r ch5}
y <- pullAncestors(x = hier, 
              node_name = deepest,
              table = T)
y
```

The attribute parameter is used to query the data in `pullAncestors` and can refer to any built in attributes of the [data.tree node network](https://cran.r-project.org/web/packages/data.tree/data.tree.pdf) or custom metadata.
```{r ch6}
y <- pullAncestors(x = hier, 
              node_name = deepest,
              table = T,
              attribute = "name")
y
```


```{r ch7}
v <- pullAncestors(x = hier, 
              node_name = deepest,
              vector = T,
              attribute = "name")
v
```

Get descendants in a similar way as above.
```{r ch8}
getDescendants(hier, v[3], attribute = "name")

```

