---
title: "CCT Hierarchy Functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CCT Hierarchy Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Load from Cache, Update if New Lineages Found or Create Cache if Missing

- Cache is written to a basic 'rda' file in the user's local AppData directory
- Will not overwrite cache if update has no new lineages compared to cache
- The force argument forcibly returns the lineages in the file specified regardless of how it compares to the cache
- Can supply either `lineages_url` or `aliases_url` to target specific files with `updateHierarchies()`
- checkCache returns a table of lineage hierarchies by default or a data.tree network with `network = TRUE`

### Table

```{r}
library(CCT)
# Grab an older version
cachedat <- checkCache(lineages_url = "https://raw.githubusercontent.com/cov-lineages/pango-designation/4b5957e151c10e92cb49e9df904c0b2ea841f3eb/lineage_notes.txt")

head(cachedat)
```


```{r}
# how many lineages?
nrow(cachedat)
```


```{r}
# Grab latest version
cachedat <- checkCache()

# how many lineages?
nrow(cachedat)
```

### Network
```{r}
cachedatNET <- checkCache(network = TRUE)
```


## Run All in One Line
- Building this documentation at a time when the master file is broken so linking specifically to my own

```{r}
hierarchy <-
  CCT::updateHierarchies(lineages_url = "https://raw.githubusercontent.com/TheZetner/pango-designation/master/lineage_notes.txt") %>% CCT::fromDFtoNetwork()

hierarchy
```

## Update Hierarchies

- Lineages: https://github.com/cov-lineages/pango-designation/blob/master/lineage_notes.txt
- Aliases: https://github.com/cov-lineages/pango-designation/blob/master/pango_designation/alias_key.json
- Can supply a specific local file or version of the files above
- Reports which file failed if one does
- Returns a table of Parent/Lineage/Children
```{r}
l <- CCT::updateHierarchies(lineages_url = "https://raw.githubusercontent.com/TheZetner/pango-designation/master/lineage_notes.txt")

head(l)
```

## From Dataframe to Network
- Take the dataframe from `updateHierarchies()` or `checkCache(network = FALSE)` and build a node network using [data.tree](https://cran.r-project.org/web/packages/data.tree/vignettes/data.tree.html)
- Defines the parent-child relationship between each node (lineage)
- Also returns a table of "orphan" lineages that are missing parents in the table as these break the network and are removed
- Can replace `ll$network` with `cachedatNET` from above
```{r}
ll <- fromDFtoNetwork(l)
ll
```

## Pull all the Ancestors of a Node

### Network
```{r}
ancestors <- pullAncestors(ll$network, "BA.5.2", table = FALSE)
# ancestors <- pullAncestors(cachedatNET, "BA.5.2", table = FALSE)

ancestors
```

### Table
```{r}
ancestors <- pullAncestors(ll$network, "BA.5.2", table = TRUE)
ancestors
```

### Vector
```{r}
# vector supersedes table param
ancestors <- pullAncestors(ll$network, "BA.5.2", table = TRUE, vector = TRUE)
ancestors
```


## Get all the Descendants of a Node

### Network
```{r}
getDescendants(ll$network, "BF.7", table = FALSE) %>% plot()
```

### Table
```{r}
getDescendants(ll$network, "BF.7", table = TRUE)
```

### Vector
```{r}
# vector supersedes table param
getDescendants(ll$network, "BF.7", table = TRUE, vector = TRUE)
```


## Make YAML Watchlist (for Augur)
```{r}
getDescendants(ll$network, "BF.7", vector = TRUE) %>% makeWatchlist()
```


## Download Variants table from CoVariants
```{r}
vars <- getVariants()
vars
```


## Expand Variants Table to Include all Children
```{r}
expandChildren(head(vars), ll$network)
```
